---
title: "a2_task2_Zoe_DuerksenSalm"
author: "Zoe Duerksen-Salm"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)

# Attach packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(dplyr)
library(lubridate)
library(patchwork)
library(broom)
library(jtools)
library(GGally)
```

With the palmetto data, use binary logistic regression to test feasibility of using variables **plant height (height)**, **canopy length (length)**, **canopy width (width)**, and **number of green leaves (green_lvs)** to classify whether a palmetto is species Serenoa repens or Sabal etonia. Use code folding and hide all messages & warnings in your knitted HTML.

```{r}
# Read in data
palmetto <- read_csv("palmetto.csv") %>% 
  clean_names()

palmetto$species <- gsub("1", "Serenoa repens", palmetto$species)
palmetto$species <- gsub("2", "Sabal etonia", palmetto$species)

palmetto$species <- as.factor(palmetto$species) 

palmetto_explore <- palmetto

palmetto_explore %>% 
  select(species, height:green_lvs) %>% 
  ggpairs(aes(color = species))
```

```{r}
ggplot(data = palmetto, aes(x = height, y = height)) +
  geom_point(aes(color = width,
                 size = green_lvs)) +
  facet_wrap(~species)
```
```{r}


palmetto_blr <- glm(species ~ height + length + width + green_lvs,   
                            data = palmetto, 
                            family = "binomial")
```


Look at the model: 
```{r}
palmetto_blr

summary(palmetto_blr)

# Get a tidy version w/ broom: 
palmetto_blr_tidy <- broom::tidy(palmetto_blr)
```

How can we start thinking about this? 

- These are coefficients for the log-linear model (e.g. these are coefficients for the predictor variables that relate to the *log odds* of the "Serenoa repens" outcome).

- The null hypothesis for coefficients is that they = 0

- The coefficient for height, `r round(palmetto_blr_tidy[2,2], 3)`, indicates that on average we expect the log odds of the species being a Serenoa repens (remember, that's the '1' outcome) increases by `r round(palmetto_blr_tidy[2,2], 3)` for each 1m increase in penguin height

Does this align with the height comparisons for the two species? 
```{r}
ggplot(data = palmetto, aes(x = species, y = height)) +
  geom_jitter()
```


- The coefficient for length, `r round(palmetto_blr_tidy[3,2], 3)`, indicates that on average we expect the log odds of the species being a Serenoa repens decreases by `r round(palmetto_blr_tidy[3,2], 3)` for each 1 mm increase in length

Does this align with the length comparisons for the two species? 
```{r}
ggplot(data = palmetto, aes(x = species, y = length)) +
  geom_jitter()
```

- The coefficient for width, `r round(palmetto_blr_tidy[4,2], 3)`, indicates that on average we expect the log odds of the species being a Serenoa repens decreases by `r round(palmetto_blr_tidy[4,2], 3)` for each 1 mm increase in width

```{r}
ggplot(data = palmetto, aes(x = species, y = width)) +
  geom_jitter()
```

```{r}
ggplot(data = palmetto, aes(x = species, y = green_lvs)) +
  geom_jitter()
```

---

But log odds are challenging to interpret. Let's find actual *probabilities* associated with a penguin being Adelie or Chinstrap, based on the selected variables and the model outcome. 

Adding `type.predict = "response"` here converts the log odds (link), the default reported, to the probability of being Chinstrap for each observation.

```{r}
palmetto_blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response")
```


Let's do a couple of quick visualizations, with green_lvs on the x-axis and probability of being a Chinstrap on the y-axis:
```{r}
# For height:
effect_plot(palmetto_blr, 
            pred = height, 
            interval = TRUE, 
            y.label = "Probability of 'Chinstrap'")

# For length:
effect_plot(palmetto_blr, 
            pred = length, 
            interval = TRUE, 
            y.label = "Probability of 'Chinstrap'")

# For width:
effect_plot(palmetto_blr, 
            pred = width, 
            interval = TRUE, 
            y.label = "Probability of 'Chinstrap'")

# For # green leavess:
effect_plot(palmetto_blr, 
            pred = green_lvs, 
            interval = TRUE, 
            y.label = "Probability of 'Chinstrap'")
```


#### Predictions for new values with `predict()`

What is the probability that a female penguin weight 3410 g with a flipper length of 192 mm will be Chinstrap?

```{r}
ex_1 <- predict(adelie_chinstrap_blr, 
                data.frame(sex = "female", 
                body_mass_g = 3410, 
                flipper_length_mm = 192),
                type = "response")

# Based on the model, the probability that this penguin is a Chinstrap is 0.4.
```

You can also feed in a new data frame, with multiple penguin observations, to get model probability estimates for more than one penguin: 

```{r}
new_df <- data.frame(
  sex = c("male", "male", "female"),
  body_mass_g = c(3298, 4100, 3600),
  flipper_length_mm = c(212, 175, 180)
)

ex_2 <- predict(adelie_chinstrap_blr,
                new_df,
                type = "response")
```
